A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-03-07 sections/introduction/main.tex
Modify introduction section

% Introduction + motivation
Emergency departments (EDs) are under increasing pressure to meet patient
waiting time targets and satisfy 
regulations~\cite{EmergencyDepartmentWinterPressures}.
It is widely reported (e.g.~\cite{mirror, thenews, bmj}) that ED congestion 
severely impacts not only patients in the ED but also Emergency Medical 
Services (EMS).
A major concern for ambulances is that they are held waiting parked outside the
ED to offload (dispatch) their patient when the ED is particularly 
busy~\cite{clarey2014ambulance}. 
Since the patient waiting time in ED is measured from the time they enter the
ED itself, there is no incentive, should the patient be stable in the
ambulance, to offload them from EMS to ED services.
As a result, ambulance blocking not only impacts on patients waiting for ED
service, but has a major knock-on effect to delaying the ability of ambulances
to respond to new EMS calls, thus placing lives at risk~\cite{eastanglia}.
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.

% ED + EMS application
The system defined in section \ref{sec:model_overview} can be applied 
to a healthcare scenario with two such queueing systems.
The two queueing systems can be thought of as two Emergency Departments (ED) 
and the Emergency Medical Services (EMS) as a distributor that distributes 
patients to them.

@@

The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and the patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts patients even if it is relatively 
full, while low threshold means that the ED blocks ambulances more frequently.

% Major contributions
In the United Kingdom the National Heath Services (NHS) sets some regulations 
on ED performance.
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

The novelty of our research contribution in this paper is to consider:
\begin{itemize}
    \item The emergence of gaming behaviour between EDs and the EMS.
    \item A queueing model with 2 consecutive waiting spaces where one would 
    serve as a parking space for the ambulances
\end{itemize}
Specifically, our focus is on the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model is then used to explore the emergent dynamics between 
the three players.
Whilst giving some consideration to equilibrium behaviour similarly to the 
work of~\cite{FirmCompetition, FirmCompetition2, deo2011centralized, 
veltman2005equilibrium}, emergent behaviour is more precisely addressed by 
considering a number of learning algorithms.
This study explores two new concepts: getting performance measures for a new
queueing theoretic model with a tandem buffer and a single service centre and 
using a learning algorithm to model the emergence of behaviour.
The developed theoretical model is then illustrated through the application to 
a healthcare system of two EDs and the EMS, exploring the inefficiencies that 
emerge and ways to apply some incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre. 
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model with a pair 
of parameters for the buffer has been considered.
% TODO Be sure if this is correct
Although our research is motivated by the particular EMS-ED example, our 
developed modelling framework and behavioural insights has application to 
similar systems across a range of sectors and settings.

% Sign posting
This paper first gives an overview of the game theoretic model, then goes on 
to describe a novel queueing model, and then follows with a number of theoretic 
results pertaining to the performance measures of this model. 
These are used to build the utilities of the game.
Finally, there is a section that describes how this game theoretic framework
can be applied at the interface between the EMS and EDs together with some 
analysis on the behaviour that is observed.

>>>>

% Introduction + motivation
Emergency departments (EDs) are under increasing pressure to meet patient
waiting time targets and satisfy 
regulations~\cite{EmergencyDepartmentWinterPressures}.
It is widely reported (e.g.~\cite{mirror, thenews, bmj}) that ED congestion 
severely impacts not only patients in the ED but also Emergency Medical 
Services (EMS).
A major concern for ambulances is that they are held waiting parked outside the
ED to offload (dispatch) their patient when the ED is particularly 
busy~\cite{clarey2014ambulance}. 
Since the patient waiting time in ED is measured from the time they enter the
ED itself, there is no incentive, should the patient be stable in the
ambulance, to offload them from EMS to ED services.
As a result, ambulance blocking not only impacts on patients waiting for ED
service, but has a major knock-on effect to delaying the ability of ambulances
to respond to new EMS calls, thus placing lives at risk~\cite{eastanglia}.

@@

% ED + EMS application
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.
The model describes the situation where an ambulance service would
have to distribute its patients between two EDs.
The two EDs can be thought of as two queueing systems and the EMS as a 
distributor that distributes patients to them, aiming to minimise some
performance measure.
The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and its patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts ambulance patients even if it is 
relatively full, while low threshold means that the ED blocks ambulances more 
frequently.
In the United Kingdom the National Heath Services (NHS) sets some regulations 
on ED performance.
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

% Major contributions
The novelty of our research contribution in this paper is to consider:
\begin{itemize}
    \item The emergence of gaming behaviour between EDs and the EMS.
    \item A queueing model with 2 consecutive waiting spaces where one would 
    serve as a parking space for the ambulances
\end{itemize}
Specifically, our focus is on the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model is then used to explore the emergent dynamics between 
the three players.
Whilst giving some consideration to equilibrium behaviour similarly to the 
work of~\cite{FirmCompetition, FirmCompetition2, deo2011centralized, 
veltman2005equilibrium}, emergent behaviour is more precisely addressed by 
considering a number of learning algorithms.
This study explores two new concepts: getting performance measures for a new
queueing theoretic model with a tandem buffer and a single service centre and 
using a learning algorithm to model the emergence of behaviour.
The developed theoretical model is then illustrated through the application to 
a healthcare system of two EDs and the EMS, exploring the inefficiencies that 
emerge and ways to apply some incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre. 
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model with a pair 
of parameters for the buffer has been considered.
Although our research is motivated by the particular EMS-ED example, our 
developed modelling framework and behavioural insights has application to 
similar systems across a range of sectors and settings.

% Sign posting
This paper consists of two main sections.
The first section presents a novel queueing model for a hospital with two
types of patients and two waiting zones.
A detailed description of how to acquire the performance measure formulas of 
such queueing system is given.
The next section gives an overview of the game theoretic model and several
theoretic results pertaining to the performance measures of this model which 
are used to build the utilities of the game.
